---
title: "A Collection of New Features"
author: "Philipp Spachtholz"
output:
  html_document:
    fig_height: 4
    fig_width: 7
    theme: paper
---

DISCLAIMER: This script was taken from [this kaggle kernel](https://www.kaggle.com/philippsp/a-collection-of-new-features/code), most of this script is not the work of the authors of this repo.

The script was reduced to add the set of engineered features which apply to both the test and the train set.

There are already a lot of good ideas for feature engineering around for this competition. However, they are all scattered across
different kernels and discussion threads.**Therefore, my plan here is to provide an easily accessible and extensible list of ideas for feature engineering (together with their implementation in R).**

```
Recent updates:
- added Feature Importance Analysis
- added more Data cleaning
- grouped appartments in the same building complex together
```

Feel free to add new ideas in the comments and
I'll add them to the list and please don't forget to **upvote** if you find it useful :-).

Having said all that, let's jump right into it:

### Initialization
```{r message=FALSE, warning=FALSE}
  library(data.table)
  library(dplyr)
  library(lubridate)
  library(stringr)
  library(caret)
  library(Matrix)
  library(geosphere)

  train <- fread('../data/test_clean.csv')
  
```

##### Some data cleaning

```{r }
train <- train %>% 
  mutate(max_floor = as.numeric(max_floor), kitch_sq=as.numeric(kitch_sq), num_room=as.numeric(num_room), build_year=as.numeric(build_year), sub_area=as.factor(sub_area)) 

train <- train %>% 
  filter(build_year < 2020 | is.na(build_year))

train <- train %>% mutate(strange_full_sq = ifelse(full_sq <= 1, full_sq+1,0), full_sq = ifelse(full_sq > 800 | full_sq <= 1, NA, full_sq))

train <- train %>% mutate(strange_life_sq = ifelse(life_sq <= 1, life_sq+1,0), strange_life_sq= ifelse(is.na(strange_life_sq),0,strange_life_sq), life_sq = ifelse(life_sq > 400 | life_sq <= 1, NA, life_sq))

train <- train %>% mutate(kitch_sq = as.numeric(kitch_sq),strange_kitch_sq = ifelse(kitch_sq <= 1, kitch_sq+1,0),kitch_sq = ifelse(kitch_sq > 200 | kitch_sq <= 1, NA, kitch_sq))


train <- train %>% mutate(num_room = as.numeric(num_room))

train <- train %>% mutate(build_year = as.numeric(build_year), strange_build_year = ifelse(build_year <= 1, build_year+1,0), build_year = ifelse(build_year > 2018 | build_year < 1860, NA, build_year))

train <- train %>% mutate(floor = ifelse(floor > 45, NA, floor))

train <- train %>% mutate(max_floor = as.numeric(max_floor), strange_max_floor = ifelse(max_floor <= 1, max_floor+1,0), max_floor = ifelse(max_floor > 60 | max_floor <=1, NA, max_floor))

train <- train %>% mutate(state = as.numeric(state), state = ifelse(state > 4, NA, state))

train <- train %>% mutate(material = as.factor(material), material = ifelse(material == 3, NA, material))

train <- train %>% mutate(product_type = factor(product_type))

train <- train %>% mutate(sub_area = factor(sub_area))

# more cleaning
train <- train %>% filter(kitch_sq < full_sq | is.na(kitch_sq))
train <- train %>% filter(kitch_sq < life_sq | is.na(kitch_sq))

train <- train %>% mutate(num_room = ifelse(num_room==0,NA,num_room))


```

### Time features

#### Split the timestamp feature

```{r }
  # year
  train <- train %>% 
    mutate(year = year(timestamp))

  # year_month
  train <- train %>% 
    mutate(year_month = make_date(year(timestamp),month(timestamp)))
```


#### Get month, week, weekday

```{r }
  # month of year
  train <- train %>% 
    mutate(month_of_year = month(train$timestamp))

  # week of year
  train <- train %>% 
    mutate(week_of_year = week(train$timestamp))

  # day of month
  train <- train %>% 
    mutate(day_of_month = mday(train$timestamp))
  
  # weekday
  train <- train %>% 
    mutate(day_of_week = wday(train$timestamp))

```


### House characteristics


```{r }
  # number of floors to the top of house
  train <- train %>% 
    mutate(floor_from_top = max_floor - floor)
  
  # relative position of floor in house
  train <- train %>% 
    mutate(floor_by_maxfloor = floor/max_floor)
  
  # average room size
  train <- train %>% 
    mutate(roomsize = (life_sq-kitch_sq)/num_room) 
  
  # relative proportion of living area
  train <- train %>% 
    mutate(life_proportion = life_sq/full_sq)
  
  # relative proportion of kitchen area
  train <- train %>% 
    mutate(kitchen_proportion = kitch_sq/full_sq)
  
  # extra area
  train <- train %>% 
    mutate(extra_area = full_sq - life_sq)
  
  # age of house at time of sale
  train <- train %>% 
    mutate(age_at_sale = interval(make_date(year=build_year),timestamp) / years(1))  
  
```

### Grouping appartments together
Some appartments seem to be in the same building. They have the same sub_area and distance to metro.
```{r}
  # assign a common name to them
  train <- train %>% 
    mutate(appartment_name = factor(str_c(sub_area,format(metro_km_avto,digits=3))))
  
  # get the number of appartments in group
  train <- train %>% 
    group_by(appartment_name) %>% 
    tally() %>% 
    right_join(train,by="appartment_name")

```



### General sale characteristics
```{r }
  # number of sales per year_month
  train <- train %>% 
    group_by(year_month) %>% 
    summarize(n_sales_permonth = n()) %>% 
    right_join(train,by="year_month")
```

#### Population characteristics
```{r }
  # population density per raion (i.e. Number of people per square meter)
  train <- train %>% 
    mutate(pop_density_raion = raion_popul/area_m)
  
  # Demographic structure of the raions
  train <- train %>% 
    mutate(young_proportion = young_all/full_all) # proportion of people younger than working age
  train <- train %>% 
    mutate(work_proportion = work_all/full_all) # proportion of people in working age
  train <- train %>% 
    mutate(retire_proportion = ekder_all/full_all) # proportion of people older than working age

```

#### Building information
```{r }
  # average building height per raion
  train <- train %>% 
    group_by(sub_area) %>% 
    summarize(mean_building_height = mean(max_floor,na.rm=T)) %>% 
    right_join(train,by="sub_area")

  # Proportion of houses with certain build materials (e.g. 10% build_count_wood)
  train<-train %>% 
    mutate_each(funs(pct = (.)/raion_build_count_with_builddate_info),matches('^build_count_[a-zA-Z]*$'))
```

#### Educational characteristics
```{r }
  # ratio of number of pupils and preschool seats
  train <- train %>% 
    mutate(ratio_preschool = children_preschool / preschool_quota)

  # ratio of number of pupils and school seats
  train <- train %>% 
    mutate(ratio_school = children_school / school_quota)

```


### dataset related features
```{r}
  # number of missing values per row (this is going to take a while)
  train <- train %>% 
    mutate(count_na_perrow = apply(., 1, function(x) sum(is.na(x))))
```


### output
```{r}
write.csv(train, file = "../data/cleanEngineeredTest.csv")
```

### Current To-do List

1. Sales (by year or month) for grouped appartments.


#### Disclaimer
1. Not all of these features are going to help your model prediction. So choose yours wisely.
2. For consistency, I will demonstrate all these features for the train set only, while for some features it would make sense to calculate them for the full dataset
3. For easier readibility, I do only minimal data cleaning in this kernel. As the new features are e.g. combinations of the basic features, it would make sense to properly clean those first.
4. While I added some features on my own, most of the features were developed by someone else. As it is not easy to figure out who had the idea first, unfortunately I cannot give you proper credt. So for all of those who shared their ideas: my gratitude goes to you. 